A generalized definition of multidimensional item response theory parameters

### 2024-06-04 output/paper_draft/paper_M2PLM_parameters.Rmd
Updates "covariance-based" version of the indices

Applying Proposition 2, if `r ANGLE_VEC_TS_ITEM` is
the vector of direction angles of `r DISCR_VECTOR`,
in the test space $`r TEST_SPACE`$ we get

@@

On the other hand, we can account for the covariance structure of
the latent space letting $`r LATENT_SPACE`$ be its geometrical representation.
Thus, we must define $`r LS_BASIS`$ in such a way that

\begin{equation}
  `r TRAIT_MV_STD_EQ`,
\end{equation}

where $`r COV_MATRIX_STD`$ is a diagonal matrix.
Now we may consider two different possibilities:
We may define $`r TRANSFORM_MATRIX`$ such that $`r INNER_PROD_CORR_COND`$.
This results in a *correlation-based* version of the indices
(note that, in this case, $`r DIAG_MATRIX_INNER_PROD_INV`$ is
an identity matrix):

\begin{equation}
  \begin{split}
    `r MDISC_CORR_PARAM_EQ` \\
    `r MIL_CORR_PARAM_EQ`.
  \end{split}
\end{equation}

However, we may also define $`r TRANSFORM_MATRIX`$ such that
$`r INNER_PROD_COV_COND`$, so

\begin{equation}
  `r TRAIT_MV_STD_NORM_EQ`.
\end{equation}

In this case, $`r DIAG_MATRIX_INNER_PROD_INV_VAR_EQ`$, resulting in the
following *covariance-based* version of the indices:

@@

>>>>

Applying Proposition 2, if $`r ANGLE_VEC_TS_ITEM`$ is
the vector of direction angles of $`r DISCR_VECTOR`$,
in the test space $`r TEST_SPACE`$ we get

@@

On the other hand, we can let the geometrical representation of
the latent space $`r LATENT_SPACE`$ account for its covariance structure,
thus giving $`r LS_BASIS`$ a statistical meaning
in the context of our MIRT modeling approach.
To do that, we may assume that $`r TRAIT_COORDS`$ is
a random vector distributed with covariance $`r COV_MATRIX_EQ`$,
being $`r COV_MATRIX`$ an $`r N_DIMS`$-dimensional positive definite matrix,
$`r CORR_MATRIX`$ its corresponding correlation matrix
(also positive definite) and $`r SD_MATRIX`$ a scaling diagonal matrix
with $`r VAR_COV_EQ`$ (i.e., the variances).

To make the representation of $`r LATENT_SPACE`$ meaningful,
the axes in $`r LS_BASIS`$ must have
an interpretation in terms of the latent space structure.
Although we do not know, in principle, what
the interpretation of a latent space basis should be, we may assume that
it must be somewhat related to the covariance structure.
Therefore, an orthonormal basis should represent independent coordinates;
formally, if $`r COV_MATRIX_ORTH`$ is the covariance matrix of
$`r TRAIT_ORTH_COORDS`$, then $`r COV_MATRIX_ORTH_EQ`$.
Applying Equation \@ref(eq:change-basis), $`r COV_MATRIX_TRANSF_EQ`$, and thus,

\begin{equation}
  \begin{split}
    `r COV_MATRIX_TRANSF_ID_EQ` \\
    `r COV_MATRIX_RESULT_EQ`.
  \end{split}
\end{equation}

Therefore, we may define $`r TRANSFORM_MATRIX`$ such that
$`r INNER_PROD_COV_COND`$;
in this case, $`r DIAG_MATRIX_INNER_PROD_INV_VAR_EQ`$, resulting in a
*covariance-based* version of the indices:

@@

```{r libraries}
#| warning: false
```
